%!TEX root = ../../Algorithms.tex
\documentclass[Chapter04]{subfiles}

\begin{document}
	\subsection{The maximum-subarray problem}

	\begin{enumerate}
		% 4.1-1
		\item What does \textsc{Find-Maximum-Subarray} return when all elements of $A$ are negative?
		\begin{answer}
			The singleton subarray $A[i \dots i]$ where $A[i] = \max_{j}\{A[j]\}$. This is due to the base case returning a singleton array.
		\end{answer}

		% 4.1-2
		\item Write pseudocode for the brute-force method of solving the maximum-subarray problem. Your procedure should run in $\Theta(n^2)$ time.
		\begin{answer}
			See \cref{alg:ch04-max-subarray-bruteforce}.

			\begin{algorithm}[H]
				\caption{A brute-force method of solving the maximum-subarray problem.}
				\label{alg:ch04-max-subarray-bruteforce}
				\SetKwProg{Fn}{def}{\string:}{end}
				\SetKwFunction{FMax}{Maximum-Subarray}
				\SetKw{KwAnd}{and}

				\Fn{\FMax{$A[1 \dots n]$}}{
					$sum_{max} \leftarrow -\infty$\;
					\For{$i \leftarrow 1\ \KwTo\ n$}{
						$sum \leftarrow 0$\;
						\For{$j \leftarrow i\ \KwTo\ n$}{
							$sum \leftarrow sum + A[j]$\;
							\If{$sum > sum_{max}$}{
								$sum_{max} \leftarrow sum$\;
								$i_{max} \leftarrow i$\;
								$j_{max} \leftarrow j$\;
							}
						}
					}
					\Return{$(i_{max}, j_{max}, sum_{max})$}
				}
			\end{algorithm}
			Our running time should be
			\begin{align*}
				\sum_{i = 1}^n \sum_{j = i}^n c &= \sum_{i = 1}^n c(n + 1 - i)\\
					&= cn^2 + cn - \frac{c}{2}n(n+1)\\
					&= \Theta(n^2).
			\end{align*}
		\end{answer}

		% 4.1-3
		\item Implement both the brute-force and recursive algorithms for the maximum-subarray problem on your own computer. What problem size $n_0$ gives the crossover point at which the recursive algorithm beats the brute-force algorithm? Then, change the base case of the recursive algorithm to use the brute-force algorithm whenever the problem size is less than $n_0$. Does that change the crossover point?
		\begin{answer}
			
		\end{answer}

		% 4.1-4
		\item Suppose we change the definition of the maximum-subarray problem to allow the result to be an empty subarray, where the sum of the values of an empty subarray is $0$. How would you change any of the algorithms that do not allow empty subarrays to permit an empty subarray to be the result?
		\begin{answer}
			Before returning the result, we can check if the sum is negative. If so, simply return an empty sub-array and a sum of 0.
		\end{answer}

		% 4.1-5
		\item Use the following ideas to develop a nonrecursive, linear-time algorithm for the maximum-subarray problem. Start at the left end of the array, and progress toward the right, keeping track of the maximum subarray seen so far. Knowing a maximum subarray of $A[1 \dots j]$, extend the answer to find a maximum subarray ending at index $j + 1$ by using the following observation: a maximum subarray of $A[1 \dots j + 1]$ is either a maximum subarray of $A[1 \dots j]$ or a subarray $A[i \dots j + 1]$, for some $1 \leq i \leq j + 1$. Determine a maximum subarray of the form $A[i \dots j + 1]$ in constant time based on knowing a maximum subarray ending at index $j$.
		\begin{answer}
			See \cref{alg:ch04-max-subarray-linear}. The algorithm uses the fact that a subarray ending at index $j + 1$ is either the singleton subarray at $j + 1$ or a subarray ending at $j$ concatenated with $j + 1$. Then the maximum subarray is either the singleton subarray or the maximum subarray ending at $j$ concatenated with $j + 1$. The singleton subarray is greater only when the maximum subarray ending at $j$ sums to a negative number.

			\begin{algorithm}[H]
				\caption{A nonrecursive, linear-time algorithm for the maximum-subarray problem.}
				\label{alg:ch04-max-subarray-linear}
				\SetKwProg{Fn}{def}{\string:}{end}
				\SetKwFunction{FMax}{Linear-Maximum-Subarray}
				\SetKw{KwAnd}{and}

				\Fn{\FMax{$A[1 \dots n]$}}{
					$sum_{max} \leftarrow -\infty$\;
					$sum_{end} \leftarrow -\infty$\;
					\For{$j \leftarrow 1\ \KwTo\ n$}{
						\eIf{$sum_{end} < 0$}{
							$i \leftarrow j$\;
							$sum_{end} \leftarrow A[j]$\;
						}{
							$sum_{end} \leftarrow sum_{end} + A[j]$\;
						}
						\If{$sum_{end} > sum_{max}$}{
							$sum_{max} \leftarrow sum_{end}$\;
							$i_{max} = i$\;
							$j_{max} = j$\;
						}
					}
				}
			\end{algorithm}
		\end{answer}

	\end{enumerate}

\end{document}